Put Title here

### 2018-11-02 General WLD combinatorics.tex
fixing the short propagator configuration lemma

\note{rank probably isn't a good word; do we have a better one?}
\begin{dfn}\label{vertex rank}
Let $i$ be a vertex in $W$. We define the {\bf rank} of $i$ to be $|Prop(i)|$, i.e. the number of propagators supported on $i$.
\end{dfn}

@@

\begin{dfn}
Define the restriction of $W$ to a region $[i,j]$ to be the subdiagram obtained by forgetting all propagators not wholly supported on vertices in the cyclic interval $[i,j]$.  Denote this diagram by $W_{[i,j]}$.
\end{dfn}

@@

\begin{lem}\label{lem sian}
  Let $W$ be an admissible WLD in which each vertex supports at least one propagator.  Then exactly one of the following things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length 3, say $p =  (i,i+3)$, with a propagator of length 2 inside it.
    \item $W$ admits no propagators of length 3, and there exists a pair of propagators with the property that the first propagator is $(i, i+1, i+2, i+3)$, the second is $(i+2, i+3, i+4, i+5)$, and no other propagator ends on the edge $(i+2, i+3)$.
  \end{enumerate}
\end{lem}

\note{This is the proof of the old result pasted without modification so we can identify where it needs editing. For some reason the WLD was named $D$ instead of $W$.}
\begin{proof}
We will prove the following stronger result: that $D$ admits a pair of propagators of length 2 in the following configuration:
\begin{equation}\label{eq:2 prop config}
\begin{tikzpicture}[baseline=(current bounding box.east)]
\pgfmathsetmacro{\radius}{2}
\pgfmathsetmacro{\angle}{20}

\begin{scope}
\clip (-\radius*1.1,\radius*0.25) rectangle (\radius*1.1,\radius*1.25);

\begin{scope}
\clip (-\radius*1.1,\radius*0.45) rectangle (\radius*1.1,\radius*1.1);
\draw[-] (0,0) circle[radius=\radius cm];
\end{scope}

% vertices
\foreach \i in {2,3,...,7}{
	\draw (\angle*\i:\radius*0.98) --  (\angle*\i:\radius*1.02);
}

\begin{scope}
\clip (0,0) circle[radius=\radius];
\draw[propagator] (2.5*\angle:\radius*1.1 cm) to [bend left=90] (4.35*\angle:\radius*1.1 cm);
\draw[propagator] (4.65*\angle:\radius*1.1 cm) to [bend left=90] (6.5*\angle:\radius*1.1 cm);

\node at (80:\radius*1.1) {\tiny $j$};
\node at (100:\radius*1.1) {\tiny $j+1$};
\node at (90:\radius*0.5) {\small $s_j = s_{j+1} = 2$};

\end{scope}
\end{tikzpicture}
\end{equation}

We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of a configuration \eqref{eq:2 prop config} in $D$ or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item $q_r$ bounds a region of $D$ which contains strictly fewer vertices than that bounded by $q_{r-1}$ and the restriction of $D$ to this region does not contain any of the propagators $p_1, \dots, p_r$.
\end{itemize}
Since $D$ contains finitely many vertices, this must eventually terminate in a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1$ of length 2 in $D$ (which exists by Remark~\ref{rem:props of length 2 and 3}), and write $(j_1,j_1+1,j_1+2,j_1+3)$ for its support.  If it is part of a configuration \eqref{eq:2 prop config} we are done, so suppose otherwise: since $s_{j_1} \geq 2$, there must exist a second propagator $q_1$ of length $\geq 4$ which starts on edge $(j_1,j_1+1)$.  This bounds a region $[i_1,j_1+1]$ which does not contain $p_1$.

Now suppose $q_{r}$ exists by the induction hypothesis and bounds a region $[i_r,j_r+1]$.  Consider the restricted diagram $D_r:=D_{[i_r,j_r+1]}$; clearly we must have $s_l \geq 2$ for all $l \in [i_r+2,j_r-1]$ (which is a non-empty interval since $\ell(q_r) \geq 4$), while $s_l\geq 1$ for $l \in \{i_r,i_r+1,j_r,j_r+1\}$.

By Remark~\ref{rem:props of length 2 and 3}, $D_r$ admits at least one propagator of length 2 and none of length 3.  Let $p_{r+1}$ be a propagator of length 2 in $D_r$; if it forms part of the configuration \eqref{eq:2 prop config} then we are done.  Otherwise, at least one endpoint of $p_{r+1}$ lies on an edge contained in the non-empty interval $[i_r+2,j_r-1]$, and so there is at least one other propagator connected to this edge.  Up to symmetry, there are three possibilities:

>>>>

\note{I don't think rank of a vertex (in the sense of number of propagators supported on it) was ever used, so I removed the definition}
%\note{rank probably isn't a good word; do we have a better one?}
%\begin{dfn}\label{vertex rank}
%Let $i$ be a vertex in $W$. We define the {\bf rank} of $i$ to be $|Prop(i)|$, i.e. the number of propagators supported on $i$.
%\end{dfn}

@@

\note{Are the next few definitions used outside the proof of Lemma~\ref{lem sian}?  If not, then perhaps they could go inside the proof so as not to clutter the document with temporary definitions?}
\begin{dfn}
Define the restriction of $W$ to a region $[i,j]$ to be the subdiagram obtained by forgetting all propagators not wholly supported on vertices in the cyclic interval $[i,j]$.  Denote this diagram by $W_{[i,j]}$.
\end{dfn}
\note{I think there's an issue here about what we mean by subdiagram; in def 1.5 we say a subdiagram is of the form $(P, V(P))$ which is certainly not what we want here; in fact I think we want the subdiagram where the vertex set is unchanged and only the propagator set is restricted (the other possibility would be to restrict the vertex set to $[i, j]$, but then the lengths of propagators can change and in the way we want to use it (see paragraph below) $p$ will become very non-admissible).  I think def 1.5 should be removed and we should say that for a WLD $(P, V)$ subdiagrams are any $(P', V')$ with $P'\subseteq P$, $V'\subseteq V$ and all supporting vertices of elements of $P'$ in $V'$; this seems to me to be the natural notion of subobject.}

@@

\begin{lem}\label{lem sian}
  Let $W$ be an admissible WLD in which each vertex supports at least one propagator.  Then exactly one of the following two things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length 3 with a propagator of length 2 inside it.
    \item $W$ admits no propagators of length 3, and one or both of the following two things occurs.
      \begin{enumerate}
        \item There is a propagator of length $2$, say $(i, i+1, i+2, i+3)$ such that no other propagators ends on either the edge $(i, i+1)$ or the edge $(i+2, i+3)$.\label{item lonely 2}
        \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+1, i+2, i+3)$, the second is $(i+2, i+3, i+4, i+5)$, and no other propagator ends on the edge $(i+2, i+3)$.\label{item pair of 2s}
      \end{enumerate}
  \end{enumerate}
\end{lem}

\begin{proof}
  Suppose first that $W$ has a propagator of length $3$, $p=(i, i+1, i+3, i+4)$.  By Remark~\ref{rm:props of length 2 and 3} along with the fact that every vertex of $W$ supports at least one propagator, we have that $i+1$ supports exactly one propagator and this propagator must have length 2 by noncrossingness.

  Now suppose $W$ has no propagators of length $3$.
%  We will prove the following stronger result: that $D$ admits a pair of propagators of length 2 in the following configuration:
%\begin{equation}\label{eq:2 prop config}
%\begin{tikzpicture}[baseline=(current bounding box.east)]
%\pgfmathsetmacro{\radius}{2}
%\pgfmathsetmacro{\angle}{20}
%
%\begin{scope}
%\clip (-\radius*1.1,\radius*0.25) rectangle (\radius*1.1,\radius*1.25);
%
%\begin{scope}
%\clip (-\radius*1.1,\radius*0.45) rectangle (\radius*1.1,\radius*1.1);
%\draw[-] (0,0) circle[radius=\radius cm];
%\end{scope}
%
%% vertices
%\foreach \i in {2,3,...,7}{
%	\draw (\angle*\i:\radius*0.98) --  (\angle*\i:\radius*1.02);
%}
%
%\begin{scope}
%\clip (0,0) circle[radius=\radius];
%\draw[propagator] (2.5*\angle:\radius*1.1 cm) to [bend left=90] (4.35*\angle:\radius*1.1 cm);
%\draw[propagator] (4.65*\angle:\radius*1.1 cm) to [bend left=90] (6.5*\angle:\radius*1.1 cm);
%
%\end{scope}
%
%\node at (80:\radius*1.1) {\tiny $j$};
%\node at (100:\radius*1.1) {\tiny $j+1$};
%\node at (90:\radius*0.5) {\small $s_j = s_{j+1} = 2$};
%
%\end{scope}
%\end{tikzpicture}
%\end{equation}

  We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of configuration \ref{item lonely 2} or \ref{item pair of 2s} from the statement
  %a configuration \eqref{eq:2 prop config} in $D$
  or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item $q_r$ bounds a region of $D$ which contains strictly fewer vertices than that bounded by $q_{r-1}$ and the restriction of $D$ to this region does not contain any of the propagators $p_1, \dots, p_r$.
\end{itemize}
Since $D$ contains finitely many vertices, this must eventually terminate in one of the two desired configurations. %a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}), and write $(j_1,j_1+1,j_1+2,j_1+3)$ for its support.  If it is part of one of the two configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise: since $p_1$ is not in configuration \ref{item lonely 2} then there must be a second propagator starting either on edge $(j_1, j_1+1)$ or $(j_1+2, j_1+3)$.  Flipping the diagram if necessary, assume the former.  Since $p_1$ is not in configuration \ref{item pair of 2s} then there must exist a second propagator $q_1$ of length $\geq 4$ which starts on edge $(j_1,j_1+1)$.  This bounds a region $[i_1,j_1+1]$ which does not contain $p_1$.

Now suppose $q_{r}$ exists by the induction hypothesis and bounds a region $[i_r,j_r+1]$.  Consider the restricted diagram $W_r:=W_{[i_r,j_r+1]}$; clearly we must have $s_l \geq 2$ for all $l \in [i_r+2,j_r-1]$ (which is a non-empty interval since $\ell(q_r) \geq 4$), while $s_l\geq 1$ for $l \in \{i_r,i_r+1,j_r,j_r+1\}$.

By Remark~\ref{rem:props of length 2 and 3}, $W_r$ admits at least one propagator of length 2 and none of length 3.  Let $p_{r+1}$ be a propagator of length 2 in $D_r$; if it forms part of either configuration \ref{item lonely 2} or \ref{item pair of 2s} 
%the configuration \eqref{eq:2 prop config}
then we are done.  Otherwise, at least one endpoint of $p_{r+1}$ lies on an edge contained in the non-empty interval $[i_r+2,j_r-1]$, and so there is at least one other propagator connected to this edge.  Up to symmetry, there are three possibilities:
